---
title: "Homework 4"
author: "Xinyi Lin"
date: "2/26/2019"
output:
  pdf_document: default
  html_document: default
---

```{r, include=FALSE}
library(tidyverse)
library(nnet)
library(MASS)
```

## Question 1

```{r, include=FALSE}
# contact low
65+130+67+54+76+48+100+111+62
# contact high
34+141+130+47+116+105+100+191+104
# tower block
65+34+54+47+100+100
# apartment
130+141+76+116+111+191
# house
67+130+48+105+62+104
```

1) contact with other residents

Contact with others/Satisfaction|Low(%)                            |Medium(%)                         |High(%)
--------------------------------|----------------------------------|----------------------------------|---------------------
Low                             |`r round((65+130+67)*100/713, 3)` |`r round((54+76+48)*100/713, 3)`  |`r round((100+111+62)*100/713, 3)`
High                            |`r round((34+141+130)*100/968, 3)`|`r round((47+116+105)*100/968, 3)`|`r round((100+191+104)*100/968, 3)`

In the medium and high satisfaction level, percentages of high contact level are larger than percentages of low contact level, which means people who contact with other residents regularly are more likely to be satisfied with housing conditions.

2) type of housing

Housing type/Satisfaction|Low(%)                         |Medium(%)                     |High(%)
-------------------------|-------------------------------|------------------------------|---------------------------
Tower block              |`r round((65+34)*100/400, 3)`  |`r round((54+47)*100/400, 3)` |`r round((100+100)*100/400, 3)`
Apartment                |`r round((130+141)*100/765, 3)`|`r round((76+116)*100/765, 3)`|`r round((111+191)*100/765, 3)`
House                    |`r round((67+130)*100/516, 3)` |`r round((48+105)*100/516, 3)`|`r round((62+104)*100/516, 3)`

Comparing to apartment and house, a higher percentage of people who live in tower block have a high satisfaction level, which means people who live in tower block are more likely to be satisfied with housing conditions. While among people who have low satisfaction, the percentage of people who live in house are largest, which means people living in house are more likely to be unsatisfied with housing conditions. 

## Question 2

### Fit model

To fit a nominal logistic regression, we first choose the category "low" as the reference ctegory and the model is:
$$\log(\frac{\pi_2}{\pi_1}) = \beta_{02} + \beta_{12}x_1 + \beta_{22}x_2 + \beta_{32}x_3$$
$$\log(\frac{\pi_3}{\pi_1}) = \beta_{03} + \beta_{13}x_1 + \beta_{23}x_2 + \beta_{33}x_3$$
in which 

$x_1$ is the indicator of contact frequency and $x_1$ = 0 means low contact level while $x_1$ = 1 means high contact level;

$x_2$ is the indicator of housing type house;

$x_3$ is the indicator of housing type tower block.

```{r, include=FALSE}
satisfaction_a = rep("sat_low", 6)
satisfaction_b = rep("sat_medium", 6)
satisfaction_c = rep("sat_high", 6)
contact = rep(c(rep("low", 3), rep("high", 3)), 3)
housing = rep(c("towerblock", "apartment", "house"), 6)

house_data = data_frame(satisfaction = c(satisfaction_a, satisfaction_b, satisfaction_c), contact = contact, housing = housing, number = c(65, 130, 67, 34, 141, 130, 54, 76, 48, 47, 116, 105, 100, 111, 62, 100, 191, 104))

house_data_nominal = 
  house_data %>% 
  spread(key = satisfaction, value = number) %>% 
  mutate(contact = factor(contact, levels = c("low", "high"))) %>% 
  dplyr::select(contact, housing, sat_low, sat_medium, sat_high)
```

```{r, include=FALSE}
sat_low = house_data_nominal$sat_low
sat_medium = house_data_nominal$sat_medium
sat_high = house_data_nominal$sat_high
contact = house_data_nominal$contact
housing = house_data_nominal$housing
house.mult <- multinom(cbind(sat_low, sat_medium, sat_high) ~ contact + housing, data=house_data_nominal)
summary(house.mult)
```

By using R, we get fitted model:

$$\log(\frac{\pi_2}{\pi_1}) = -0.514 + 0.296x_1 + 0.070x_2 + 0.407x_3$$
$$\log(\frac{\pi_3}{\pi_1}) = -0.081 + 0.328x_1 + -0.304x_2 + 0.642x_3$$

### Odd ratios with 95% CI

The fitted model we get:

```{r, echo=FALSE}
summary(house.mult)
```

model |term        |estimated odd ratio|95% CI
------|------------|-------------------|------------------------------------------------------
medium|contact_high|`r exp(0.2960)`    |`r exp(0.2960+c(qnorm(0.025),-qnorm(0.025))*0.1301)`
medium|house       |`r exp(0.0697)`    |`r exp(0.0697+c(qnorm(0.025),-qnorm(0.025))*0.1438)`
medium|towerblock  |`r exp(0.4068)`    |`r exp(0.4068+c(qnorm(0.025),-qnorm(0.025))*0.1713)`
high  |contact_high|`r exp(0.3282)`    |`r exp(0.3282+c(qnorm(0.025),-qnorm(0.025))*0.1182)`
high  |house       |`r exp(-0.3040)`   |`r exp(-0.3040+c(qnorm(0.025),-qnorm(0.025))*0.1352)`
high  |towerblock  |`r exp(0.6416)`    |`r exp(0.6416+c(qnorm(0.025),-qnorm(0.025))*0.1501)`

### The pattern in the associations

According to estimated odd ratios, we can find that in medium and high models, estimated odd ratios of contact high vs low are larger than 1, which means that people with high contact level are more likely to have higher satisfaction level. We can also find that in medium and high models, estimated odd ratios of towerblock vs apartment are larger than 1, which means that comparing with people living in apartment, people living in tower block are more likely to have higher satisfaction level. And this pattern meet what we find in previous tables.

### Test the goodness of fit

Now, we need to test the goodness of fit.

```{r, include=FALSE}
pihat=predict(house.mult,type='probs') 
m=rowSums(house_data_nominal[,3:5])
res.pearson=(house_data_nominal[,3:5]-pihat*m)/sqrt(pihat*m)
G.stat=sum(res.pearson^2) # Generalized Pearson Chisq Stat
G.stat
pval=1-pchisq(G.stat,df=(6-4)*(3-1)) # fit is good
pval
# deviance
D.stat=sum(2*house_data_nominal[,3:5]*log(house_data_nominal[,3:5]/(m*pihat)))
D.stat
pval=1-pchisq(D.stat,df=(6-4)*(3-1))
pval
```

By using R to calculate, we get generalized Pearson $\chi^2$ statistic:$$G = \sum_{i=1}^n\sum_{j=1}^J\frac{(y_{ij}-m_i\hat\pi_{ij})^2}{m_i\hat\pi_{ij}} = \sum_{i=1}^n\sum_{j=1}^JR_{p_{ij}}^2 = 6.932$$ As it follows $\chi^2(4)$, corresponding p-value is 0.14 which is larger than 0.05, so this model fit the data well.

## Question 3

We use following proportional odds model to fit data:

$$\log(\frac{\pi_1}{\pi_2+\pi_3}) = \beta_{01} + \beta_1x_1 + \beta_2x_2 + \beta_3x_3$$
$$\log(\frac{\pi_1+\pi_2}{\pi_3}) = \beta_{02} + \beta_1x_1 + \beta_2x_2 + \beta_3x_3$$
in which 

$x_1$ is the indicator of contact frequency and $x_1$ = 0 means low contact level while $x_1$ = 1 means high contact level;

$x_2$ is the indicator of housing type house;

$x_3$ is the indicator of housing type tower block.

By using R, we get fitted model as follow:

```{r,include=FALSE}
house_data_ordinal = 
  house_data %>% 
  mutate(contact = factor(contact, levels = c("low", "high"))) %>% 
  mutate(satisfaction = factor(satisfaction, levels = c("sat_low", "sat_medium", "sat_high")))
```

```{r, echo=FALSE}
house.polr=polr(satisfaction~contact+housing,data=house_data_ordinal,weights=number)
summary(house.polr) # pay attention to sign, read help doc of polr (-eta)
```

which is:

$$\log(\frac{\pi_1}{\pi_2+\pi_3}) = -0.496 - 0.2524x_1 + 0.2353x_2 - 0.5010x_3$$
$$\log(\frac{\pi_1+\pi_2}{\pi_3}) = 0.616 - 0.2524x_1 + 0.2353x_2 - 0.5010x_3$$

According to results, as $\beta_1=-0.2524$, people who have high contact level are more likely to have higher satisfaction level given house types are same. As $\beta_2 = 0.2353$, comparing to people who live in apartment, people who live in house are more likely to have lower satisfaction level given contact levels are same. As $\beta_3 = -0.5010$, comparing to people who live in apartment, people who live in tower block are more likely to have higher satisfaction level given contact levels are same.

## Question 4

By using R, we get Pearson residuals as follow:

```{r, echo=FALSE}
pihat=predict(house.polr,house_data_nominal,type='p')
m=rowSums(cbind(house_data_nominal$sat_low, house_data_nominal$sat_medium, house_data_nominal$sat_high))
res.pearson=(house_data_nominal[,3:5]-pihat*m)/sqrt(pihat*m)
res.pearson
```

According to results, we can find the largest discrepancy is the discrepancy of frequency in people who live in house with high contact level and high satisfaction level and the discrepancy is -1.478. 