---
title: "Homework 5"
author: "Xinyi Lin"
date: "3/5/2019"
output:
  pdf_document: default
  html_document: default
---

```{r, results='hide'}
library(readr)
```

## Problem 1

### Question 1

```{r, message=FALSE}
# import data
crab_data <- read_table2("./HW5-crab.txt")
```

Using W as the single predictor, we can fit following model:$$log(\lambda_i) = \beta_0 + \beta_1x_i$$ where $x_i$ represents carapace width(W).

By using R, we fit the Posisson model(M1) as follow:
```{r}
m1.glm = glm(Sa~W, family=poisson, data=crab_data)
summary(m1.glm)
```

which is:$$log(\lambda_i) = -3.305 + 0.164x_i$$

```{r, results='hide'}
res.p1=residuals(m1.glm,type='pearson',data=crab_data)
G1.stat=sum(res.p1^2)
G1.stat
df = 173-2
pval = 1-pchisq(G1.stat,df=df) # chisq test
pval
```

By using R, we know that the generalize Pearson $\chi^2$ equals to 544.157 and corresponding p-value equals to 0, which means this model doesn't fit the data.

Interpretation:
$\beta_1 = 0.164$ means when the width of female crab increases one unit, the log-number of satellites increases 0.164.

### Question 2

Using both W and Wt as  predictors, we can fit following model:$$log(\lambda_i) = \beta_0 + \beta_1x_{1i} + \beta_2x_{2i}$$ where $x_1$ represents carapace width(W) and $x_2$ represents carapace weight(Wt).

By using R, we fit the Posisson model(M2) as follow:

```{r}
m2.glm = glm(Sa~W + Wt, family=poisson, data=crab_data)
summary(m2.glm)
```

which is: $$log(\lambda_i) = -1.292 - 0.046x_{1i} + 0.447x_{2i}$$

We use deviance to compare two models and null hypothesis is that two model have similar performance.

```{r, results='hide'}
test.stat=m1.glm$deviance-m2.glm$deviance
df=171-170
pval=1-pchisq(test.stat,df=df) # chisq test
pval
```

By using R, we can get p-value equals to $0.00469$, which is smaller than 0.05, so we reject null hypothesis and conclude that bigger model(M2) performs far more better.

Interpretation:

$\beta_1 = -0.046$ means that with one unit increase in width, the log-number of satellites for one female crab decreases 0.046 given the weight keeps the same.;

$\beta_2 = 0.447$ means that with one unit increase in weight, the log-number ratio of satellites for one female crab increases 0.447 given the width keeps the same.

### Question 3

First, we need to calculate the goodness of fit of M2.

```{r, results='hide'}
res.p2=residuals(m2.glm,type='pearson',data=crab_data)
G2.stat=sum(res.p2^2)
G2.stat
df = 173-3
pval = 1-pchisq(G2.stat,df=df) # chisq test
pval
```

By using R, we know that the generalize Pearson $\chi^2$ equals to 536.596, corresponding p-value equals to 0, which means this model doesn't fit the data and there might be over dispersion.

```{r, results='hide'}
phi=G2.stat/(173-3)
phi 
m2.glm$deviance/m2.glm$df.residual
```

By using R, we get dispersion parameter $\phi = 3.156$ and half normal plot is as follow:

```{r}
plot(qnorm((173+1:173+0.5)/(2*173+1.125)),sort(abs(res.p2)),
     xlab='Expected Half-Normal Order Stats',
     ylab='Ordered Abs Pearson Residuals')
abline(a=0,b=1)
abline(a=0,b=sqrt(phi),lty=2)
```

By using R, we get adjusted model as follow:

```{r}
summary(m2.glm,dispersion=phi)
```

which is: $$log(\lambda_i) = -1.292 - 0.046x_{1i} + 0.447x_{2i}$$

Interpretation:

$\beta_1 = -0.046$ means that with one unit increase in width, the log-number of satellites for one female crab decreases 0.046 given the weight keeps the same.;

$\beta_2 = 0.447$ means that with one unit increase in weight, the log-number ratio of satellites for one female crab increases 0.447 given the width keeps the same.

Estimated values of parameters are the same in the adjusted model, while variances increase.

## Problem 2

### Question 1

```{r, message=FALSE, warning=FALSE}
# import data
parasite_data <- read_table2("./HW5-parasite.txt") 
parasite_data = na.omit(parasite_data)
parasite_data$Year = as.factor(parasite_data$Year)
parasite_data$Area = as.factor(parasite_data$Area)
```

With predictors area, year and length, we can fit following model:$$log(\lambda_i) = \beta_0 + \beta_1x_{1i} + \beta_2x_{2i} + \beta_3x_{3i} + \beta_4x_{4i} + \beta_5x_{5i} + \beta_6x_{6i}$$ where $x_1$ is a indicator of area2, $x_2$ is a indicator of area3, $x_3$ is a indicator of area4, $x_4$ is a indicator of year 2000, $x_5$ is a indicator of year 2001, $x_6$ is a indicator of length.

```{r}
parasite.glm = glm(Intensity ~ Area + Year + Length, family=poisson, data=parasite_data)
summary(parasite.glm)
```

which is:$$log(\lambda_i) = 2.643 - 0.212x_{1i} - 0.117x_{2i} + 1.405x_{3i} + 0.6703x_{4i} - 0.2181x_{5i} - 0.0284x_{6i}$$

Interpretation:

$\beta_1 = -0.212$ means comparing to area 1, the log-number of parasites in fish living in area 2 decreases 0.212 given year and length are the same.

$\beta_2 = -0.117$ means comparing to area 1, the log-number of parasites in fish living in area 3 decreases 0.117 given year and length are the same.

$\beta_3 = 1.405$ means comparing to area 1, the log-number of parasites in fish living in area 4 increases 1.405 given year and length are the same.

$\beta_4 = 0.6703$ means comparing to year 1999, the log-number of parasites in fish increases 0.6703 in year 2000 given area and length are the same.

$\beta_5 = -0.2181$ means comparing to year 1999, the log-number of parasites in fish decrease 0.2181 in year 2001 given area and length are the same.

$\beta_6 = -0.0284$ means with one unit increases in length, the log-number of parasites in fish decrease 0.0284 given area and year are the same.

### Question 2

```{r, results='hide'}
res.p=residuals(parasite.glm,type='pearson',data=parasite_data)
G.stat=sum(res.p^2)
G.stat
df = 1191-6
pval = 1-pchisq(G2.stat,df=df) # chisq test
pval
```

By using R, we know that the generalize Pearson $\chi^2$ equals to 42164.97, corresponding p-value equals to 0, which means this model doesn't fit the data.

### Question 3

We assume whether a fish is susceptible to parasites depends on area, year and length and how many parasites in one fish depends on area, year and length.

Let $Z_i$ be a latent binary variable that generates structural zeros $$P(Z_i = 0) = \pi_i$$. 

The response satisfies $$Y_i|(Z_i = 0) = 0$$ $$Y_i|(Z_i = 1) \sim Poisson(\lambda_i)$$

Then we get corresponding models: $$\log(\frac{\pi_i}{1-\pi_i}) = \alpha_0 + \alpha_1x_{1i} + \alpha_2x_{2i} + \alpha_3x_{3i} + \alpha_4x_{4i} + \alpha_5x_{5i} + \alpha_6x_{6i}$$ $$\log(\lambda_i) = \beta_0 + \beta_1x_{1i} + \beta_2x_{2i} + \beta_3x_{3i} + \beta_4x_{4i} + \beta_5x_{5i} + \beta_6x_{6i}$$ where $x_1$ is a indicator of area2, $x_2$ is a indicator of area3, $x_3$ is a indicator of area4, $x_4$ is a indicator of year 2000, $x_5$ is a indicator of year 2001, $x_6$ is a indicator of length.

We use R and fit following models:

```{r, message=FALSE}
library(pscl)
zero.model <- zeroinfl(Intensity ~ Area + Year + Length, data = parasite_data) 
summary(zero.model)
```

which are:
$$\log(\frac{\pi_i}{1-\pi_i}) = 0.553 + 0.719x_{1i} + 0.658x_{2i} - 1.023x_{3i} - 0.752x_{4i} + 0.457x_{5i} - 0.010x_{6i}$$ $$\log(\lambda_i) = 3.843 + 0.269x_{1i} + 0.146x_{2i} + 0.945x_{3i} + 0.392x_{4i} - 0.045x_{5i} - 0.037x_{6i}$$

Interpretation:

$\alpha_0 = 0.553$ means the log odd of a fish is not susceptible to parasites is 0.553;

$\alpha_1 = 0.719$ means the log odd ratio of a fish is not susceptible to parasites in area 2 vs area 1 is 0.719 given length and year are the same;

$\alpha_2 = 0.658$ means the log odd ratio of a fish is not susceptible to parasites in area 3 vs area 1 is 0.658 given length and year are the same;

$\alpha_3 = 1.023$ means the log odd ratio of a fish is not susceptible to parasites in area 4 vs area 1 is 1.023 given length and year are the same;

$\alpha_4 = -0.752$ means the log odd ratio of a fish is not susceptible to parasites in year 2000 vs year 1999 is -0.752 given area and length are the same;

$\alpha_5 = 0.457$ means the log odd ratio of a fish is not susceptible to parasites in year 2001 vs year 1999 is 0.457 given area and length are the same;

$\alpha_6 = -0.010$ means the log odd ratio of a fish is not susceptible to parasites is 0.658 with one unit increases in length given area and year are the same.

$\beta_1 = 0.269$ means comparing to area 1, the log-number of parasites in fish living in area 2 increases 0.269 given year and length are the same when fishes are susceptible to parasites;

$\beta_2 = 0.146$ means comparing to area 1, the log-number of parasites in fish living in area 3 increases 0.146 given year and length are the same when fishes are susceptible to parasites;

$\beta_3 = 0.945$ means comparing to area 1, the log-number of parasites in fish living in area 4 increases 0.945 given year and length are the same when fishes are susceptible to parasites;

$\beta_4 = 0.392$ means comparing to year 1999, the log-number of parasites in fish increases 0.392 in year 2000 given area and length are the same when fishes are susceptible to parasites;

$\beta_5 = -0.045$ means comparing to year 1999, the log-number of parasites in fish decrease 0.045 in year 2001 given area and length are the same when fishes are susceptible to parasites;

$\beta_6 = -0.037$ means with one unit increases in length, the log-number of parasites in fish decrease 0.037 given area and year are the same when fishes are susceptible to parasites.

## Appendix Code
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```