---
title: "Homework 8"
author: "Xinyi Lin"
date: "4/21/2019"
output:
  pdf_document: default
  html_document: default
---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(gee)
library(lme4)
```

```{r}
library(readxl)
health <- read_excel("HEALTH.xlsx") %>% 
  janitor::clean_names() %>% 
  mutate(health = as.factor(health))
```

## Question 1

Draw a plot about numbers of 'Good' health group and 'Poor' health group in control group and intervention group when the time equals to 1.

```{r}
health %>% 
  filter(time == 1) %>% 
  ggplot(aes(x=txt, fill=health)) +
  geom_bar(position="dodge") + 
  xlab("txt") + labs(fill="health")
```

According to the plot, ratios of people whose self-rating levels are good or poor separetly in control group are slightly different from those in intervention group.

Fit a GLM model to find out whether group assignment is a significant variable when predicting self-rated level.

```{r}
health_random = health %>% 
  filter(time == 1)
glm1 = glm(health~txt, family = binomial, health_random)
summary(glm1)
```

According to the model, we can find that `txt` variable is not significant so there is no significant relationship between randomized group assignment and participants health self-rating.

## Question 2

Fit the GEE model with unstructure correlation structure.

```{r, message=FALSE, warning=FALSE}
subset <- subset(health, time > 1)
# make month 0 as another covariate: baseline
base = subset(health, time == 1) %>% 
  select(id, health)
colnames(base) = c("id", "baseline")
new_subset = merge(subset,base,by="id",all=F)
new_subset$nhealth <- as.numeric(new_subset$health == "Good") # 1=Good, 0=Poor
new_subset$time = plyr::mapvalues(new_subset$time, from = c(2,3,4), to = c(3,6,12))

#head(new_subset)

gee.fit = gee(nhealth~txt+time+agegroup+baseline, 
                 data = new_subset, family = "binomial", 
                 id = id ,corstr = "unstructured", 
                 scale.fix = TRUE, scale.value = 1) 
summary(gee.fit)
```

Interpretation: 

Intercept: The log odd of self-rating of health status as good is 0.12 on average for sub-population in control, age 15-24 group at the time of randomization with Good baseline.

$\beta_{txt}:$ The log odd ratio of self-rating of health status as good between intervention group vs control group is 2.10 on average for sub-population with same age group, baseline in same time.

$\beta_{time}:$ The log odd ratio of self-rating of health status as good is 0.03 on average with one unit changes in time for sub-population with same age group, baseline and treatment group.

$\beta_{age25-34}:$ The log odd ratio of self-rating of health status as good between 25-34 age group vs 15-24 age group is 1.35 on average for sub-population with same treatment group, baseline in same time.

$\beta_{age35+}:$ The log odd ratio of self-rating of health status as good between 35+ age group vs 15-24 age group is 1.42 on average for sub-population with same treatment group, baseline in same time.

$\beta_{base}:$ The log odd ratio of self-rating of health status as good between poor baseline group vs good baseline group is -1.81 on average for sub-population with same treatment group and age group in same time.

## Question 3

Fit GLMM model. The model is as following:

$$logit(E(Y_{ij}|b_i)) = (b_i+\beta_1) + X_{ij}^T\beta$$

```{r}
glmm.fit <- glmer(nhealth ~ baseline + txt + time + agegroup + (1 | id), 
                  family = 'binomial', data = new_subset)
summary(glmm.fit)
```

Interpretation:

$\beta_0:$ The log odd of self-rating of health status as good is 0.20 on average across all subjects with baseline equals to good, in control group and 15-24 age group at the time of randomization.

$\beta_{base}:$ Cannot interpret.

$\beta_{txt}:$ Cannot interpret.

$\beta_{time}:$ The log odd ration of self-rating of health status as good is 0.04 with one unit changes in time for the same subject.

$\beta_{age25-34}:$ Cannot interpret.

$\beta_{age35+}:$ Cannot interpret.

Difference:

For the GEE model, it considers the situation of sub-population, so its interpretation is based on sub-population and all coefficients can be interpreted. However, for the GLMM mdoel, it considers specific individual and some of its coefficients.