---
title: "Homework 9"
author: "Xinyi Lin"
date: "4/25/2019"
output:
  pdf_document: default
  html_document: default
---

## Problem 1

As $$h(x) = \frac{2x}{1+x^2} = \frac{d[ln(1+x^2)]}{dx} = -\frac{d[-ln(1+x^2)]}{dx}$$ $$\log S(x) = -ln(1+x^2)$$ survival function is $$S(x) = e^{-ln(1+x^2)} = \frac{1}{1+x^2}$$
$$F(x) = 1-S(x) = \frac{x^2}{1+x^2}$$
so the density function is $$f(x) = \frac{dF(x)}{dx} = \frac{2x}{(1+x^2)^2}$$

## Problem 2

Let $\hat{F(t)}$ be the Fleming-Harrington estimator of $S(t)$. The Kaplan-Meier estimate, the Fleming-Harrington estimate of survival function and the Nelson-Aalen estimate of the cumulative hazard function are shown below.

\begin{tabular}{cccccccc}
\hline
$t_i$ & $n_i$ & $d_i$ & $c_i$ & $\hat \lambda_i$ & $\hat{S_t}$ & $\tilde{H_t}$ & $\hat{F_t}$\\
\hline
1 & 10 & 1 & 0 & 1/10 & 0.90 & 0.10 & 0.90\\
2 & 9 & 2 & 0 & 2/9 & 0.70 & 0.32 & 0.72\\
4 & 7 & 0 & 1 & 0/7 & 0.70 & 0.32 & 0.72\\
5 & 6 & 0 & 1 & 0/6 & 0.70 & 0.32 & 0.72\\
6 & 5 & 1 & 0 & 1/5 & 0.56 & 0.52 & 0.59\\
7 & 4 & 0 & 1 & 0/4 & 0.56 & 0.52 & 0.59\\
8 & 3 & 0 & 1 & 0/3 & 0.56 & 0.52 & 0.59\\
9 & 2 & 0 & 1 & 0/2 & 0.56 & 0.52 & 0.59\\
10 & 1 & 0 & 1 & 0/1 & 0.56 & 0.52 & 0.59\\
\hline
\end{tabular}

## Problem 3

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(survival)
```

```{r}
library(KMsurv)
data(tongue)
#head(tongue)
```

### Aneuploid Tumor

```{r}
aneu_data = tongue %>% 
  filter(type == 1)
```

The fitted model and K-M curve of aneuploid tumor are as following:

```{r}
KM1=survfit(Surv(time,delta)~1, data = aneu_data, conf.type='log')
plot(KM1, mark.time = TRUE,xlab="Weeks", ylab="Survival Probability", 
     main="K-M curve", cex.lab=1.5, cex.main=1.5)
#plot(KM1)
#plot(KM1,fun='cumhaz') # cumulative hazard fun
```

Its pointwise 95% confidence interval is:

```{r}
summary(KM1)[9:10]
```

There are 52 weeks ony year, so the estimated 1-year survival rate and 95% CI are as following:

```{r}
summary(KM1, times = 52)
```

So the estimated survival rate is 0.654 and 95% CI is (0.537, 0.797).

### Diploid Tumor

```{r}
dipl_data = tongue %>% 
  filter(type == 2)
```

The fitted model and K-M curve of aneuploid tumor are as following:

```{r}
KM2=survfit(Surv(time,delta)~1, data = dipl_data, conf.type='log')
plot(KM2, mark.time = TRUE,xlab="Weeks", ylab="Survival Probability", 
     main="K-M curve", cex.lab=1.5, cex.main=1.5)
#plot(KM2)
#plot(KM2,fun='cumhaz') # cumulative hazard fun
```

Its pointwise 95% confidence interval is:

```{r}
summary(KM2)[9:10]
```

There are 52 weeks ony year, so the estimated 1-year survival rate and 95% CI are as following:

```{r}
summary(KM2, times = 52)
```

So the estimated survival rate is 0.486 and 95% CI is (0.33, 0.716).